S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light

### 2017-04-05 s-aligner.tex
Reformat text to make it easier to view/edit.

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The availability and amount of sequenced genomes have been rapidly growing in recent years due to the adoption of next-generation sequencing (NGS) technologies that allow for high-throughput short-read generation at highly competitive cost. Since this trend is expected to continue in the foreseeable future, the design and implementation of efficient and scalable NGS bioinformatics algorithms are important to research and industrial applications.  In this paper, we introduce S-Aligner---a highly scalable read mapper designed for the \emph{Sunway Taihu Light} supercomputer and its fourth-generation ShenWei many-core architecture (SW26010). S-Aligner employs a combination of optimization techniques to overcome both the memory-bound and the compute-bound bottlenecks in the read mapping algorithm. In order to make full use of the compute power of Sunway Taihu Light, our design employs three levels of parallelism: (1) inter-node parallelism using MPI based on a task-grid pattern, (2) intra-node parallelism using multi-threading and asynchronous data transfer to fully utilize all 260 cores of the SW26010 many-core processor, and (3) vectorization to exploit the available 256-bit SIMD vector registers.  Moreover, we have employed asynchronous access patterns and data sharing strategies during file I/O to overcome bandwidth limitations of the network file system. Our performance evaluation demonstrates that S-Aligner scales almost linearly with approximately 95\% efficiency for up to 13,312 nodes (concurrently harnessing more than 3 million compute cores). Furthermore, our implementation on a single node outperforms the established RazerS3 mapper running on a platform with eight Intel Xeon E7-8860v3 CPUs while achieving highly competitive alignment accuracy.
\end{abstract}

@@

\subsection{Related Work}
We are not the first to parallelize read mapping on a compute cluster. Among the earliest tools are pBWA \cite{pbwa} and pMap \cite{pmap}. pBWA is an MPI implementation of Version 0.5.9 of the BWA aligner while pMap provides an MPI-based framework for the concurrent execution of popular aligners such as BWA \cite{bwa} and Bowtie \cite{bowtie}. Other implementations exploit MapReduce frameworks, for example, the Hadoop-based SEAL \cite{seal} and BigBWA \cite{bigbwa} tools or the Spark-based SparkBWA suite \cite{sparkBWA}, in order to guarantee fault tolerance. Unfortunately, all these approaches suffer from insufficient scalability when executed on hundreds of nodes. The metaframework parSRA \cite{parSRA} is designed for read mappers written in UPC++.  It combines dynamic scheduling and a virtual file system layer for read distribution to overcome the limitations imposed by pMap and MapReduce-based approaches. merAligner \cite{merAligner} and CUSHAW3-UPC++ \cite{cushaw3upc} are other PGAS-based short-read aligners written in UPC; both demonstrate good scalability. Although pMap and parSRA allow for the execution of single node GPU aligners such as nvBowtie \cite{nvBio}, none of the cited tools is explicitly designed for heterogeneous clusters consisting of thousands of many-core processors. Furthermore, all previous approaches target traditional hardware architectures.

@@

\subsection{Seed-and-Extend Approach}
%There are some "all" mappers currently on traditional commercial platforms, RazerS3\cite{razers3} is one of the most famous "all" mapper on x86\_64 CPUs. Hobbes2\cite{hobbes2}, BitMapper \cite{bitmapper} and mrFAST\cite{mrfast} are also read mappers in "all" mode on x86\_64 platforms. BitMapper2 also supports CUDA-enabled GPUs.

@@

\item[Stage 3. Alignment.] This stage generates the base-pair level alignment information of a read and its verified intervals in $G$. 
%{\color{red} Note that some mappers just use a traceback of Myers delta matrix and some use Smith-waterman, so we shouldn't mention the algorithm}
\end{description}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/GridNew}
	\caption{Our task-partitioning and file-loading strategy: (a) overall design; (b) and (c) detailed behavior of processes.}
	\label{TaskGrid}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/AsyncRead}
	\caption{Asynchronous data loading strategy: (a) default case, and (b) case when there is no significantly shorter reference block.}
	\label{AsyncRead}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/FrmWk}
	\caption{Workflow of S-Aligner on a single CG: gray boxes correspond to tasks assigned to SPs while tasks in white boxes are executed on the MP.}
	\label{FrmWk}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/AsyncPara}
	\caption{Asynchronous filtration.}
	\label{AsyncPara}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/SPParallel}
	\caption{Framework of our parallelization scheme within one CG: the MP performs filtration and output of alignment results while SPs verify identified candidate locations and generate base-pair level alignments. Gray boxes indicate the intervals verified in the second round of verification (assuming $slice\_num = 4$).}
	\label{SPParallel}
	%\end{center}
\end{figure}

@@

%In our experiments, the dual buffer strategy shows a performance improvement between 50\% to 70\% in the verification stage by minimizing the idle times of SPs. This optimization technique is also used for the alignment stage.

%\subsection{Using Intrinsics}

\subsection{SIMD Vectorization}
Implementing Myers' bit-parallel algorithm using ShenWei's SIMD intrinsics requires a bit-level encoding of DNA sequences to efficiently evaluate the characteristic function $\chi(s_{i-1} = s'_{j-1})$ for all $i, j$.

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=0.9\linewidth]{figures/MixPack}
	\caption{Bit-level encoding strategies for DNA sequences. For simplicity, we use a 4-bit word instead of 64-bit word for illustrating our combined AoSoA strategy (bottom).}
	\label{MixPack}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
    \begin{lstlisting}[frame=single, xleftmargin=3.5ex]
t1    = simd_vxorw(ref_hi,read_hi[k]);
t2    = simd_vxorw(ref_lo,read_lo[k]);
X     = simd_vbisw(t1,t2); 
X     = simd_vxorw(X,one); 
X     = simd_vbisw(X,VP[k]);
D0    = simd_vandw(X,VP[k]);
D0    = simd_uaddo_take_carry(D0,VP[k]);
D0    = simd_vandw(D0,VP[k]);
D0    = simd_vandw(D0,X);
HN    = simd_vandw(VP[k],D0);
HP    = simd_vbisw(VP[k],D0);
HP    = simd_vxorw(HP,one);
HP    = simd_vbisw(HP,VN[k]);	
X     = simd_sllow(HP,resi32);
X     = simd_vbisw(X,pr_HP);
pr_HP = simd_srlow(HP,255);
VN[k] = simd_vandw(X,D0);
VP[k] = simd_vbisw(X,D0);
VP[k] = simd_vxorw(VP[k],one);
t2    = simd_sllow(HN,resi32);
t2    = simd_vbisw(t2,pr_HN);
pr_HN = simd_srlow(HN,255);
VP[k] = simd_vbisw(t2,VP[k]);	
    \end{lstlisting}
%	%\begin{center}
%	%{\scriptsize
%	{\footnotesize
%		\begin{tabular}{| l |}\hline
%			\begin{minipage}{0.95\hsize}
%				\verb| |\\
%				\verb|  |~~1. 	t1 = simd\_vxorw(ref\_hi,read\_ hi[k]);\\
%				\verb|  |~~2.	t2 = simd\_vxorw(ref\_lo,read\_ lo[k]);\\
%				\verb|  |~~3. 	X = simd\_vbisw(t1,t2); \\
%				\verb|  |~~4.   X = simd\_vxorw(X,one); \\
%				\verb|  |~~5.  	X = simd\_vbisw(X,VP[k]); \\
%				\verb|  |~~6.  	D0 = simd\_vandw(X,VP[k]); \\
%				\verb|  |~~7.   D0 = simd\_uaddo\_take\_carry(D0,VP[k]); \\
%				\verb|  |~~8.   D0 = simd\_vandw(D0,VP[k]);\\
%				\verb|  |~~9.  	D0 = simd\_vandw(D0,X);\\
%				\verb|  |~10. 	HN = simd\_vandw(VP[k],D0); \\
%				\verb|  |~11.   HP = simd\_vbisw(VP[k],D0); \\
%				\verb|  |~12.  	HP = simd\_vxorw(HP,one);\\
%				\verb|  |~13.   HP = simd\_vbisw(HP,VN[k]);\\	
%				\verb|  |~14.	X = simd\_sllow(HP,resi32);\\
%				\verb|  |~15.   X = simd\_vbisw(X,pre\_HP);\\
%				\verb|  |~16.	pre\_HP = simd\_srlow(HP,255); \\
%				\verb|  |~17.	VN[k] = simd\_vandw(X,D0);\\
%				\verb|  |~18.	VP[k] = simd\_vbisw(X,D0); \\
%				\verb|  |~19.	VP[k] = simd\_vxorw(VP[k],one);\\
%				\verb|  |~20.	t2 = simd\_sllow(HN,resi32);\\
%				\verb|  |~21.	t2 = simd\_vbisw(t2,pre\_HN); \\
%				\verb|  |~22.	pre\_HN = simd\_srlow(HN,255);\\
%				\verb|  |~23.	VP[k] = simd\_vbisw( t2 ,VP[k]);\\	
%			\end{minipage}\\
%			\hline
%		\end{tabular}
%	}
%	%\end{center}
	\caption{SIMD core instructions used to implement Myers' algorithm on ShenWei. The reference sequence is presented as \texttt{ref\_hi}, the high-bit of the pattern, and \texttt{ref\_lo}, the low-bit of the pattern. The read sequence is stored in the vectors \texttt{read\_hi[k]} and \texttt{read\_lo[k]}. The variables \texttt{D0}, \texttt{HN}, \texttt{HP}, \texttt{VN}, \texttt{VP} refer to the corresponding variables in Equation~\ref{one-hot-myers}.}
	\label{cores}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/AsyncTrans}
	\caption{Asynchronous data transfer to LDM.}
	\label{AsyncTrans}
	%\end{center}
\end{figure}

@@

\section{Performance Evaluation}
\label{Evaluation}
The performance of S-Aligner has been evaluated on the Sunway Taihu Light supercomputer. We use GRCh38\footnote{available at http://hgdownload.cse.ucsc.\\edu/downloads.html} as the human reference genome.  For the read input data sets we use either ERR013135\footnote{available at ftp://ftp.sra.ebi.ac.uk/vol1/fastq/ERR013/\\ERR013135} or reads simulated by Mason \cite{mason} or wgsim \cite{wgsim}.\footnote{Note that Mason usually generates reads of higher quality but does not support long reads.}  We first evaluate the performance in terms of runtime and mapping accuracy on a single node. Subsequently, we evaluate the scalability of our implementation by varying the number of utilized nodes up to 13,312.

@@

\subsection{Comparison with RazerS3}
In this experiment, we use the first chromosome from GRCh38 as reference and various number of reads from ERR013135. The runtime of S-Aligner on a single SW26010 node is compared with RazerS3, a representative all-mapper. Since RazerS3 does not support ShenWei's architecture, we ran it on a machine with Intel processors using multithreading with default parameters (e.g., the accuracy parameter is set to 98\%). Measured runtimes are shown in Table \ref{SingleNode}.  One can see that S-Aligner executed on a single SW26010 (260 cores running at 1.45 GHz) outperforms RazerS3 running on eight Xeon E7-8860v3 CPUs (128 cores running at 3.2 GHz).

\begin{table}
	\begin{threeparttable}
		\caption{Runtime comparison between S-Aligner running on a single CG and RazerS3 running on a single node with eight Xeon E7-8860v3 CPUs.}
		\label{SingleNode}
		\begin{tabular}{@{\extracolsep{2pt}}rrrrr}
			\hline
			\multicolumn{1}{c}{Ref} &
			\multicolumn{2}{c}{Reads} &
			\multicolumn{2}{c}{Time(s)}\\
			\cline{2-3}
			\cline{4-5}
			\multicolumn{1}{c}{\#bps} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} &
			\multicolumn{1}{c}{RazerS3\tnote{\textdagger}} &
			\multicolumn{1}{c}{S-Aligner\tnote{\textdaggerdbl}}\\
			\hline
			%65M & 10M  & 108 &  52\tnote{\textdagger} & 47\tnote{\textdaggerdbl}\\
			%58M & 40M  & 108 &  1415 & 435\\
			116M & 40M & 108 &  939 & 405\\
			253M & 40M & 108 &  3,044 & 892\\
			%58M & 40M  & 200 &  898 & 1,002\\
			%116M & 40M & 200 &  744 & 1,013\\
			253M & 40M & 200 &  3,430 & 2,580\\
			\hline
		\end{tabular}
		\begin{tablenotes}
		\item[\textdagger] result is from a machine with eight Xeon E7-8860v3 (128 cores up to 3.20GHz).
		\item[\textdaggerdbl] result is from a node with a SW26010 (260 cores with frequency of 1.45GHz). 
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

%As we can see, our algorithm has better performance than RazerS3, with an accuracy no worse than ordinary RazerS3.

\begin{table}
	\begin{threeparttable}
		\caption{Accuracy evaluation of S-Aligner for both real and simulated data.}
		\label{AccuEval}
		\begin{tabular}{@{\extracolsep{2pt}}rrrrrr}
			\hline
			\multicolumn{1}{c}{Chrom} &
			\multicolumn{3}{c}{Reads} &
			\multicolumn{1}{c}{\multirow{2}{*}{Interv. found}} \\
%			\multicolumn{1}{c}{Normalized}\\
			\cline{2-4}
			\multicolumn{1}{c}{Index} &
			\multicolumn{1}{c}{Origin} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} \\		
%			\multicolumn{1}{c}{intervals found}}\\
			\hline
			%65M & 10M  & 108 &  52\tnote{\textdagger} & 47\tnote{\textdaggerdbl}\\
			%58M & 40M  & 108 &  1415 & 435\\
			1\textsuperscript{st} & ERR013135 & 20,000 & 108 & 99.34\%\\
			1\textsuperscript{st} & Mason &  20,000 & 200 & 99.82\%\\
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

@@

\begin{table}[!htb]
	\begin{threeparttable}
		\caption{Performance and runtime evaluation of S-Aligner using different numbers of nodes.}
		\label{paraexp}
		\begin{tabular}{rrrrrr}
			\hline
			\multicolumn{1}{c}{\multirow{2}{*}{\#nodes}} &
			\multicolumn{2}{c}{Reads}  &
			\multicolumn{1}{c}{\multirow{2}{*}{Time(s)}} &
			\multicolumn{2}{c}{Performance (bpps)} \\
			\cline{2-3}
			\cline{5-6}
			\multicolumn{1}{c}{} &
			\multicolumn{1}{c}{Count} &
			\multicolumn{1}{c}{\#bps} & &
			\multicolumn{1}{c}{Total (M)} &
			\multicolumn{1}{c}{Node (K)}\\
			\hline
			13 & 16M & 200 & 1,315 & 2.43 & 185.43\\
			%			26 & 64 & 2664 & 4.9 & 187\\
			52 & 64M & 200 & 1,321 & 9.69 & 184.80\\
			%			104 & 64 & 2664 & 4.9 & 187\\
			208 & 256M & 200 & 1,321 & 38.76 & 184.66\\
			%			416 & 64 & 2664 & 4.9 & 187\\
			832 & 1,024M & 200 & 1,327 & 154.33 & 184.59\\
			%			1664 & 64 & 2664 & 4.9 & 187\\
			3,328 & 4,096M & 200 & 1,349 & 607.26 & 179.67\\
			13,312 & 4,096M & 200 & 344 & 2,381.40 & 178.90\\
			\hline
		\end{tabular}
		\begin{tablenotes}
			\item
			\#nodes is the number of nodes involved in computing.
			\item
			bpps is short for million bp processed per second.
			\item
			M / K indicates that column is presented in millions / thousands.
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

% that's all folks
\end{document}

>>>>

\begin{abstract}
The availability and amount of sequenced genomes have been rapidly
growing in recent years due to the adoption of next-generation
sequencing (NGS) technologies that allow for high-throughput
short-read generation at highly competitive cost. Since this trend is
expected to continue in the foreseeable future, the design and
implementation of efficient and scalable NGS bioinformatics algorithms
are important to research and industrial applications.  In this paper,
we introduce S-Aligner---a highly scalable read mapper designed for
the \emph{Sunway Taihu Light} supercomputer and its fourth-generation
ShenWei many-core architecture (SW26010). S-Aligner employs a
combination of optimization techniques to overcome both the
memory-bound and the compute-bound bottlenecks in the read mapping
algorithm. In order to make full use of the compute power of Sunway
Taihu Light, our design employs three levels of parallelism: (1)
inter-node parallelism using MPI based on a task-grid pattern, (2)
intra-node parallelism using multi-threading and asynchronous data
transfer to fully utilize all 260 cores of the SW26010 many-core
processor, and (3) vectorization to exploit the available 256-bit SIMD
vector registers.  Moreover, we have employed asynchronous access
patterns and data sharing strategies during file I/O to overcome
bandwidth limitations of the network file system. Our performance
evaluation demonstrates that S-Aligner scales almost linearly with
approximately 95\% efficiency for up to 13,312 nodes (concurrently
harnessing more than 3 million compute cores). Furthermore, our
implementation on a single node outperforms the established RazerS3
mapper running on a platform with eight Intel Xeon E7-8860v3 CPUs
while achieving highly competitive alignment accuracy.
\end{abstract}

@@

\subsection{Related Work}

We are not the first to parallelize read mapping on a compute
cluster. Among the earliest tools are pBWA \cite{pbwa} and pMap
\cite{pmap}. pBWA is an MPI implementation of Version 0.5.9 of the BWA
aligner while pMap provides an MPI-based framework for the concurrent
execution of popular aligners such as BWA \cite{bwa} and Bowtie
\cite{bowtie}. Other implementations exploit MapReduce frameworks, for
example, the Hadoop-based SEAL \cite{seal} and BigBWA \cite{bigbwa}
tools or the Spark-based SparkBWA suite \cite{sparkBWA}, in order to
guarantee fault tolerance. Unfortunately, all these approaches suffer
from insufficient scalability when executed on hundreds of nodes. The
metaframework parSRA \cite{parSRA} is designed for read mappers
written in UPC++.  It combines dynamic scheduling and a virtual file
system layer for read distribution to overcome the limitations imposed
by pMap and MapReduce-based approaches. merAligner \cite{merAligner}
and CUSHAW3-UPC++ \cite{cushaw3upc} are other PGAS-based short-read
aligners written in UPC; both demonstrate good scalability. Although
pMap and parSRA allow for the execution of single node GPU aligners
such as nvBowtie \cite{nvBio}, none of the cited tools is explicitly
designed for heterogeneous clusters consisting of thousands of
many-core processors. Furthermore, all previous approaches target
traditional hardware architectures.

@@

\subsection{Seed-and-Extend Approach}

@@

\item[Stage 3. Alignment.] This stage generates the base-pair level
  alignment information of a read and its verified intervals in $G$.

\end{description}

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/GridNew}
  \caption{Our task-partitioning and file-loading strategy: (a)
    overall design; (b) and (c) detailed behavior of processes.}
  \label{TaskGrid}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/AsyncRead}
  \caption{Asynchronous data loading strategy: (a) default case, and
    (b) case when there is no significantly shorter reference block.}
  \label{AsyncRead}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/FrmWk}
  \caption{Workflow of S-Aligner on a single CG: gray boxes correspond
    to tasks assigned to SPs while tasks in white boxes are executed
    on the MP.}
  \label{FrmWk}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/AsyncPara}
  \caption{Asynchronous filtration.}
  \label{AsyncPara}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=\linewidth]{figures/SPParallel}
  \caption{Framework of our parallelization scheme within one CG: the
    MP performs filtration and output of alignment results while SPs
    verify identified candidate locations and generate base-pair level
    alignments. Gray boxes indicate the intervals verified in the
    second round of verification (assuming $slice\_num = 4$).}
  \label{SPParallel}
\end{figure}

@@

\subsection{SIMD Vectorization}

Implementing Myers' bit-parallel algorithm using ShenWei's SIMD
intrinsics requires a bit-level encoding of DNA sequences to
efficiently evaluate the characteristic function $\chi(s_{i-1} =
s'_{j-1})$ for all $i, j$.

@@

\begin{figure}[!htb]
  \includegraphics[width=0.9\linewidth]{figures/MixPack}
  \caption{Bit-level encoding strategies for DNA sequences. For
    simplicity, we use a 4-bit word instead of 64-bit word for
    illustrating our combined AoSoA strategy (bottom).}
  \label{MixPack}
\end{figure}

@@

\begin{figure}[!htb]
    \begin{lstlisting}[frame=single, xleftmargin=3.5ex]
t1    = simd_vxorw(ref_hi,read_hi[k]);
t2    = simd_vxorw(ref_lo,read_lo[k]);
X     = simd_vbisw(t1,t2); 
X     = simd_vxorw(X,one); 
X     = simd_vbisw(X,VP[k]);
D0    = simd_vandw(X,VP[k]);
D0    = simd_uaddo_take_carry(D0,VP[k]);
D0    = simd_vandw(D0,VP[k]);
D0    = simd_vandw(D0,X);
HN    = simd_vandw(VP[k],D0);
HP    = simd_vbisw(VP[k],D0);
HP    = simd_vxorw(HP,one);
HP    = simd_vbisw(HP,VN[k]);	
X     = simd_sllow(HP,resi32);
X     = simd_vbisw(X,pr_HP);
pr_HP = simd_srlow(HP,255);
VN[k] = simd_vandw(X,D0);
VP[k] = simd_vbisw(X,D0);
VP[k] = simd_vxorw(VP[k],one);
t2    = simd_sllow(HN,resi32);
t2    = simd_vbisw(t2,pr_HN);
pr_HN = simd_srlow(HN,255);
VP[k] = simd_vbisw(t2,VP[k]);	
    \end{lstlisting}
    \caption{SIMD core instructions used to implement Myers' algorithm
      on ShenWei. The reference sequence is presented as
      \texttt{ref\_hi}, the high-bit of the pattern, and
      \texttt{ref\_lo}, the low-bit of the pattern. The read sequence
      is stored in the vectors \texttt{read\_hi[k]} and
      \texttt{read\_lo[k]}. The variables \texttt{D0}, \texttt{HN},
      \texttt{HP}, \texttt{VN}, \texttt{VP} refer to the corresponding
      variables in Equation~\ref{one-hot-myers}.}
    \label{cores}
\end{figure}

@@

\begin{figure}[!htb]
  \includegraphics[width=1\linewidth]{figures/AsyncTrans}
  \caption{Asynchronous data transfer to LDM.}
  \label{AsyncTrans}
\end{figure}

@@

\section{Performance Evaluation}
\label{Evaluation}

The performance of S-Aligner has been evaluated on the Sunway Taihu
Light supercomputer. We use GRCh38\footnote{available at
  http://hgdownload.cse.ucsc.\\edu/downloads.html} as the human
reference genome.  For the read input data sets we use either
ERR013135\footnote{available at
  ftp://ftp.sra.ebi.ac.uk/vol1/fastq/ERR013/\\ERR013135} or reads
simulated by Mason \cite{mason} or wgsim \cite{wgsim}.\footnote{Note
  that Mason usually generates reads of higher quality but does not
  support long reads.}  We first evaluate the performance in terms of
runtime and mapping accuracy on a single node. Subsequently, we
evaluate the scalability of our implementation by varying the number
of utilized nodes up to 13,312.

@@

\subsection{Comparison with RazerS3}

In this experiment, we use the first chromosome from GRCh38 as
reference and various number of reads from ERR013135. The runtime of
S-Aligner on a single SW26010 node is compared with RazerS3, a
representative all-mapper. Since RazerS3 does not support ShenWei's
architecture, we ran it on a machine with Intel processors using
multithreading with default parameters (e.g., the accuracy parameter
is set to 98\%). Measured runtimes are shown in Table
\ref{SingleNode}.  One can see that S-Aligner executed on a single
SW26010 (260 cores running at 1.45 GHz) outperforms RazerS3 running on
eight Xeon E7-8860v3 CPUs (128 cores running at 3.2 GHz).

\begin{table}
  \begin{threeparttable}
    \caption{Runtime comparison between S-Aligner running on a single
      CG and RazerS3 running on a single node with eight Xeon
      E7-8860v3 CPUs.}
    \label{SingleNode}
    \begin{tabular}{@{\extracolsep{2pt}}rrrrr}
      \hline
      \multicolumn{1}{c}{Ref} &
      \multicolumn{2}{c}{Reads} &
      \multicolumn{2}{c}{Time(s)}\\
      \cline{2-3}
      \cline{4-5}
      \multicolumn{1}{c}{\#bps} &
      \multicolumn{1}{c}{Count} &
      \multicolumn{1}{c}{\#bps} &
      \multicolumn{1}{c}{RazerS3\tnote{\textdagger}} &
      \multicolumn{1}{c}{S-Aligner\tnote{\textdaggerdbl}}\\
      \hline
      116M & 40M & 108 &  939 & 405\\
      253M & 40M & 108 &  3,044 & 892\\
      253M & 40M & 200 &  3,430 & 2,580\\
      \hline
    \end{tabular}
    \begin{tablenotes}
    \item[\textdagger] result is from a machine with eight Xeon
      E7-8860v3 (128 cores up to 3.20GHz).
    \item[\textdaggerdbl] result is from a node with a SW26010 (260
      cores with frequency of 1.45GHz).
    \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

\begin{table}
  \begin{threeparttable}
    \caption{Accuracy evaluation of S-Aligner for both real and
      simulated data.}
    \label{AccuEval}
    \begin{tabular}{@{\extracolsep{2pt}}rrrrrr}
      \hline
      \multicolumn{1}{c}{Chrom} &
      \multicolumn{3}{c}{Reads} &
      \multicolumn{1}{c}{\multirow{2}{*}{Interv. found}} \\
      \cline{2-4}
      \multicolumn{1}{c}{Index} &
      \multicolumn{1}{c}{Origin} &
      \multicolumn{1}{c}{Count} &
      \multicolumn{1}{c}{\#bps} \\		
      \hline
      1\textsuperscript{st} & ERR013135 & 20,000 & 108 & 99.34\%\\
      1\textsuperscript{st} & Mason &  20,000 & 200 & 99.82\%\\
      \hline
    \end{tabular}
  \end{threeparttable}
\end{table}

@@

\begin{table}[!htb]
  \begin{threeparttable}
    \caption{Performance and runtime evaluation of S-Aligner using
      different numbers of nodes.}
    \label{paraexp}
    \begin{tabular}{rrrrrr}
      \hline
      \multicolumn{1}{c}{\multirow{2}{*}{\#nodes}} &
      \multicolumn{2}{c}{Reads}  &
      \multicolumn{1}{c}{\multirow{2}{*}{Time(s)}} &
      \multicolumn{2}{c}{Performance (bpps)} \\
      \cline{2-3}
      \cline{5-6}
      \multicolumn{1}{c}{} &
      \multicolumn{1}{c}{Count} &
      \multicolumn{1}{c}{\#bps} & &
      \multicolumn{1}{c}{Total (M)} &
      \multicolumn{1}{c}{Node (K)}\\
      \hline
      13 & 16M & 200 & 1,315 & 2.43 & 185.43\\
      52 & 64M & 200 & 1,321 & 9.69 & 184.80\\
      208 & 256M & 200 & 1,321 & 38.76 & 184.66\\
      832 & 1,024M & 200 & 1,327 & 154.33 & 184.59\\
      3,328 & 4,096M & 200 & 1,349 & 607.26 & 179.67\\
      13,312 & 4,096M & 200 & 344 & 2,381.40 & 178.90\\
      \hline
    \end{tabular}
    \begin{tablenotes}
    \item
      \#nodes is the number of nodes involved in computing.
    \item
      bpps is short for million bp processed per second.
    \item
      M / K indicates that column is presented in millions / thousands.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

\end{document}
